\ttitle

### 2013-11-22 Chapters/Chapter2.tex
'Refactoring'

\section{Overview}\label{sec:literature_review_overview}
As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.

\TODO{Give examples of temporal classification techniques}

In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.

\TODO{Include highlights from old-classification text}

In the following section of this is chapter we begin with a review of current literature on temporal segmentation.
Many techniques of temporal segmentation rely in change detection algorithms.
In \Cref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
An often encountered problem is the \emph{curse of dimensionality}.
To overcome this problem, we look at \gls{dr} methods in \Cref{sec:dimensionality_reduction}.
We end this chapter with a review on \gls{svm} based methods and algorithms.
This latter section relates the \gls{oc-svm} methods via \gls{dr} to temporal segmentation.

\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}
\input{./Chapters/Chapter2/dimensionality_reduction.tex}
\input{./Chapters/Chapter2/svm_based.tex}

>>>>

\input{./Chapters/Chapter2/overview.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}
% \input{./Chapters/Chapter2/dimensionality_reduction.tex}
\input{./Chapters/Chapter2/svm_based.tex}

### 2013-11-22 Chapters/Chapter2/cusum.tex
WIP: Working on organization of Chapter 2
'Refactoring'

% !TEX root = ../../main.tex
\section{CUSUM}\label{cusum}

@@

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
Using the \gls{icss} algorithm they are able to find multiple change points in a reflective manner.
\TODO{Explain ``reflective'', or use term as batch, post-processing etc.}
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum of squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

>>>>

% !TEX root = ../../main.tex
\section{CUSUM}\label{sec:cusum}

@@

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum of squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

### 2013-11-22 Chapters/Chapter2/temporal_segmentation.tex
'Refactoring'

@@

This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.
Finally we will discuss other applications of \glspl{svm} in the context of these terms.

\subsection{Segmentation}\label{subsec:segmentation}
\TODO{This subsection is mainly from previous draft version}

\TODO{Create compact notation, one sentence per paper max}

Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  Given the average segment length $L$, the complexity of this method is $O(nL)$.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
\end{itemize}
It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.

The \gls{swab} method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
Each segment is approximated by using piecewise linear representation (PLR), an often used method.
The user provided error threshold controls the granularity and number of segments.
Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.

@@

--- Segmentation ---

@@

Section~\ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

@@

% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\

>>>>

Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.

@@

This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.

Aspects we will consider when discussing the literature are applicability to \gls{har}, parameter tuning and ... \TODO{What more aspects?}
% Finally we will discuss other applications of \glspl{svm} in the context of these terms.

\subsection{Segmentation}\label{subsec:segmentation}

Methods that apply temporal segmentation on time series data can be roughly categorized in three different methods, as discussed by Avci \etal \cite{avci2010activity}.
In that survey the distinction between \emph{Top-Down}, \emph{Bottom-Up}, and \emph{Sliding-Window} approaches is made, based on the way data is processed by the algorithm.
Since we are interested in on-line change detection, the literature discussed in this section will mainly be forms of sliding-window algorithms.

In Keogh \etal \cite{keogh2001online} a comparison of the aforementioned approaches is made.
The research also introduces the \gls{swab} method, which combines the simple and intuitive approach of the sliding-window approach with the (semi-)global view over the data from the bottom-up method.
This proposed method is dependent on user provided parameters, setting the maximum error when performing the algorithm.
Each segment is approximated by using a \gls{plr}.
The provided error threshold controls the granularity and number of segments.

Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.

@@

% --- Segmentation ---

@@

\Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

@@

% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguita2012human}. 13, 2012 \\

### 2013-11-22 Chapters/Chapter3.tex
'Refactoring'

>>>>

\TODO{Give this a nice place in this chapter somewhere, on model building}
Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  Given the average segment length $L$, the complexity of this method is $O(nL)$.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
\end{itemize}
It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered. \\
--- end ---

\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}
